---
title: Investigate the exponential distribution in R and compare it with the Central
  Limit Theorem
author: "Rahul Deo"
date: "14/11/2021"
output:
  pdf_document: default
  html_document: default
---

## Overview
In this exercise, I shall try to compare the exponential distribution with central limit theorem keeping the parameter lambda =0.3. I choose this because its mean and variance are simple to interpret. 


## Simulations
Set the simulation variables lambda, exponentials, and seed.
```{r}
library('ggplot2')
ECHO=TRUE
set.seed(42)
lambda = 0.3
exponentials = 40
```

Run Simulations with variables
```{r}
cal_mean = NULL
for (i in 1 : 1000) cal_mean = c(cal_mean, mean(rexp(exponentials, lambda)))
```


## Sample Mean versus Theoretical Mean comaprision 
#### Sample Mean
Calculating the mean from the collected sample means.
```{r}
mean(cal_mean)
```

#### Theoretical Mean
Theoretical mean is inverse(lambda) or lambda^-1
```{r}
lambda^-1
```

#### Comparison
Ideally, there should be a slight difference bettwen the theoretical and calculated values.
```{r}
abs(mean(cal_mean)-lambda^-1)
```


## Sample Variance versus Theoretical Variance
#### Sample Variance
Calculated Variance
```{r}
var(cal_mean)
```

#### Theoretical Variance
Theoratical variance 
(lambda * sqrt(n))^-2.
```{r}
(lambda * sqrt(exponentials))^-2
```

#### Comparison
Difference between both the variances
```{r}
abs(var(cal_mean)-(lambda * sqrt(exponentials))^-2)
```


## Distribution
A histogram plot overlayed with normal distribution so that It can represent the concept that any sequence will become normal if we sample a lot of of points. The normal distribution will have a mean of lambda^-1 and standard deviation of (lambda*sqrt(n))^-1, the theoretical normal distribution for the simulations.
```{r}
library(ggplot2)
ggplot(data.frame(y=cal_mean), aes(x=y)) + 
  geom_histogram(aes(y=..density..), binwidth=0.2, fill="#0072B2", 
                 color="black") +
  geom_function(fun=dnorm, args = c(mean=lambda^-1, sd=(lambda*sqrt(exponentials))^-1),
  size=2) +
  labs(title="Plot of Calculated Means and Overlay Normal Distribution", x="Drawn sample Mean")
```
